---
title: 'Reproducing Research: Causal Effect of Government Transfers and Government Support'
date: ""
output:
 pdf_document:
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




# Empirical Analysis using Data from Manacorda, Miguel, & Vigorito (2011, American Economic Journal: Applied Economics)


This exercise uses data from Manacorda, Miguel, & Vigorito's paper, "Government Transfers and Political Support," published in the *American Economic Journal: Applied Economics* in 2011. This paper studies how receipt of a government anti-poverty cash transfer changes how beneficiary households support and view the government. 

The data can be found on Edward Miguel's faculty website. Download and extract the contents from the `Government_Transfers_replication.zip` file. 


# Set up and constructing the data

The original data used in the paper is confidential. The authors instead provide the `reg_panes.dta` data file which is anonymized and created from the original data. 


Open the `reg_panes.dta` file. To complete this problem set you will need the following variables from this data file: 

| Name            |Description                                        |
|-----------------|---------------------------------------------------|
|aprobado         |Ever received PANES 2005-2007                      |
|untracked07      | Untracked in 2007                                 |
|h_89             |Supports current government 2007 [1 to 3]|         |
|hv34             |Supports current government 2008 [1 to 3]|         | 
|ind_reest        |Predicted Income                                   | 
|newtreat         |PANES eligibility                                  |


\pagebreak

## Loading the Packages

Load any R packages we will be using:

**Code:**
```{r,warning=F,message=F}
library(haven)
library(dplyr)
library(ggplot2)
library(lfe)
library(stargazer)
library(kableExtra)
library(statar)
```










Drop all other variables. If needed, give the variables you are keeping more intuitive names. 


**Code:**

```{r}
df<- read_dta('C:\\Users\\roryq\\Downloads\\reg_panes.dta')%>%
  select(aprobado,untracked07,h_89, hv34, ind_reest,newtreat)
```


## **The data as downloaded will require that you clean the variables of interest and construct a new dataset to generate the graphs. Start by generating the following cleaned variable:**

-An indicator for receiving PANES that is NA if a respondent is untracked in 2007

**Code:**

```{r}
df$indicator_panes <- ifelse(df$untracked07 == 1, NA, df$aprobado)

```


\pagebreak

## **We are going to re-scale the variables that indicate support for the current government so that responses range from 0 to 1. To do this, tabulate the current variable to see how it is distributed and then generate a variable that will be NA if it is currently coded as 9, 0 if currently 2, 0.5 if currently 1 and 1 if currently 3. Do this for both the 2007 and 2008 variable. **

Note: This is how the authors modify this variable in their code. It seems counter intuitive and does not correspond to the description of how this variable is coded in the survey questionnaire as reported in their appendix though it does correspond to their discussion in footnote 12. My guess is the transcription/translation of the survey question is incorrect.

**Code:**

```{r}
table(df$h_89) %>% kbl() %>% kable_classic(full_width = F,
 html_font = "Cambria",latex_options = "hold_position")

table(df$hv34) %>% kbl() %>% kable_classic(full_width = F,
 html_font = "Cambria",latex_options = "hold_position")
```

```{r}
df$h_89_scaled <- ifelse(df$h_89 == 9, NA,
                         ifelse(df$h_89 == 2, 0,
                                ifelse(df$h_89 == 1, 0.5,
                                       ifelse(df$h_89 == 3, 1, NA))))

# Rescale support for the current government in 2008 (hv34)
df$hv34_scaled <- ifelse(df$hv34 == 9, NA,
                          ifelse(df$hv34 == 2, 0,
                                 ifelse(df$hv34 == 1, 0.5,
                                        ifelse(df$hv34 == 3, 1, NA))))

```

\pagebreak

## **Generate a variable that is the square of predicted income.**



**Code:**
```{r}
df$income_sq<- df$ind_reest
```


# We start by reproducing the main figures (2, 3,and 4) of the paper as good figures are key to any regression discontinuity paper.


### **The data consists of over 3000 observations. How many points are plotted on these figures? How should we interpret the y axis? How many points are plotted below the threshold? How many points are plotted above the threshold?**



**Answer:**

There are 45 points, because we will split the data into percentile groups (15 below the threshold and 30 above).  Each of these points is the average gov support, and predicted income within each of these percentile groups.


### **Why is the number of points above the threshold different from the number below? **

**Answer:**

Because many more observations fell below the threshold than those above, so to make sure each percentile consisted of the same number of observations, we needed more percentiles below, and these percentiles are what is graphed.  Leading to more obserevations plotted below than above.



### **Replicating these figures will require restructuring our data and calculating the values that are plotted. Generate a variable that will indicate the percentile group the observation is in. Note the difference in the number of percentile groups above and below the threshold. **


**Code:**

```{r,warning=F}

 
df$pct[df$ind_reest<0]<-xtile(df$ind_reest[df$ind_reest<0],n=30)
df$pct1[df$ind_reest>=0]<-xtile(df$ind_reest[df$ind_reest>=0],n=15)
df$pct1<-df$pct1+30
df$pct[df$ind_reest>=0]<-df$pct1[df$ind_reest>=0]


```


\pagebreak


## **For each of the percentile groups, calculate the mean of each of the variables we will use for plotting: predicted income, receipt of PANES, support for the government in 2007, and support for the government in 2008.**

**Code:**

```{r}
df_summary <- df %>%
  group_by(pct) %>%
  summarize(
    mean_predicted_income = mean(ind_reest, na.rm = TRUE),
    mean_indicator_panes = mean(indicator_panes, na.rm = TRUE),
    mean_support_2007 = mean(h_89_scaled, na.rm = TRUE),
    mean_support_2008 = mean(hv34_scaled, na.rm = TRUE)
  )

df_summary %>% kbl %>% kable_classic(full_width = F, html_font = "Cambria")
```


\pagebreak


## **Replicate figure 2. Make the figure as clear and informative as possible. You may want to create an indicator variable for percentiles above and below the threshold. ** 

**Code:**

```{r, warning=F}
df <- df %>%
  mutate(
    above_threshold = ifelse(ind_reest >= 0, 1, 0)
  )

# Create the plot
ggplot(df, aes(x = ind_reest, y = aprobado, color = factor(above_threshold))) +
  geom_point(alpha = 0.6, size = 2) + 
  scale_color_manual(values = c("cornflowerblue", "firebrick"),
                     labels = c("Below Threshold", "Above Threshold")) +  
  geom_vline(xintercept = 0, linetype = "dashed", color = "black") + 
  labs(
    title = "PANES Program Eligibility and Participation",
    x = "Predicted Income",
    y = "",
    color = "Eligibility Group",
     caption = "Figure 2") +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold",hjust = 0.5),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10),
    plot.caption = element_text(size = 10, face = "italic"),
    legend.position = "top"
  )
```

\pagebreak


## **What is the purpose of this figure and what should we take away from it? **

**Answer:**
  The purpose of this graph is to determine if it is a fuzzy or sharp regression continuity, by evaluating the discontinuity at the threshold between x and treatment.

  The sharp discontinuity of those that received panes falling below the threshold, show a strict enforcement of the program eligibility threshold



## **Replicate figures  3 and 4. Make these figures as clear and informative as possible.  **

```{r, message=F}
ggplot(df_summary, aes(x = mean_predicted_income, y = mean_support_2007)) + 
  geom_point(aes(fill = mean_predicted_income < 0), size = 2, alpha = 0.6, stroke = 0.5,
             shape = 21, color = "black") +  
  geom_smooth(data = subset(df_summary, mean_predicted_income < 0),
              aes(x = mean_predicted_income, y = mean_support_2007),
              method = "lm", se = FALSE, color = "cornflowerblue", alpha = 0.6) +  
  geom_smooth(data = subset(df_summary, mean_predicted_income >= 0),
              aes(x = mean_predicted_income, y = mean_support_2007),
              method = "lm", se = FALSE, color = "firebrick", alpha = 0.6) + 
  geom_vline(xintercept = 0, linetype = "dashed", color = "black") +
  labs(
    title = "PANES Program Eligibility and Political Support (2007)",
    x = "Predicted Income",
    y = "Support for Current Government",
    caption = "Figure 3."
  ) +
  scale_fill_manual(values = c("TRUE" = "cornflowerblue", "FALSE" = "firebrick"),
                    labels = c("Below Threshold", "Above Threshold"),
                    breaks = c("TRUE", "FALSE")) + 
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10),
    plot.caption = element_text(size = 10, face = "italic"),
    legend.position = "top"
  ) +
  guides(fill = guide_legend(title = ""))  


```

\pagebreak

```{r,message=F}
ggplot(df_summary, aes(x = mean_predicted_income, y = mean_support_2008)) + 
  geom_point(aes(fill = mean_predicted_income < 0), size = 2, alpha = 0.6, stroke = 0.5,
             shape = 21, color = "black") +  
  geom_smooth(data = subset(df_summary, mean_predicted_income < 0),
              aes(x = mean_predicted_income, y = mean_support_2008),
              method = "lm", se = FALSE, color = "cornflowerblue", alpha = 0.6) + 
  geom_smooth(data = subset(df_summary, mean_predicted_income >= 0),
              aes(x = mean_predicted_income, y = mean_support_2008),
              method = "lm", se = FALSE, color = "firebrick", alpha = 0.6) + 
  geom_vline(xintercept = 0, linetype = "dashed", color = "black") + 
  labs(
    title = "PANES Program Eligibility and Political Support (2008)",
    x = "Predicted Income",
    y = "Support for Current Government",
    caption = "Figure 4."
  ) +
  scale_fill_manual(values = c("TRUE" = "cornflowerblue", "FALSE" = "firebrick"),
                    labels = c("Below Threshold", "Above Threshold"),
                    breaks = c("TRUE", "FALSE")) +  
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10),
    plot.caption = element_text(size = 10, face = "italic"),
    legend.position = "top"
  ) +
  guides(fill = guide_legend(title = ""))  



```


\pagebreak


## **Interpret these figures. What should we take away from them? **

**Answer:** 

**Figure 3:** 

The estimated discontinuity implies that program eligibility was associated with a 13 percentage point increase in support for the government over the opposition coalition. This provides evidence that households’ political views are responsive to government transfers

**Figure 4:** 

similar though somewhat smaller gains in FA support—of between 8 and 12 percentage points

**Overall:** suggests that past transfers also factor meaningfully into voters’ decision making.

\pagebreak

## **Replicate the results of the three regressions estimated in the first column of table 1. Present your results in a table. Interpret the coefficients.**

**Code:**

```{r, results='asis'}
# Regression 1: Ever received PANES (2005–2007)
model1 <- lm(indicator_panes ~ newtreat, data = df)

# Regression 2: Government support (2007)
model2 <- lm(h_89_scaled ~ newtreat, data = df)

# Regression 3: Government support (2008)
model3 <- lm(hv34_scaled ~ newtreat, data = df)


stargazer(model1, model2, model3,
          type = "latex",
          dep.var.labels = c("Ever Received PANES", "Gov. Support 2007", "Gov. Support 2008"),
          covariate.labels = c("Coefficient (Standard Error)"),omit.stat = c("ser"),
          column.labels = c("(1)", "(2)", "(3)"),digits = 3)

```


**Answer:**

The coefficient of 0.993 suggests that being eligible for the PANES program almost 99.3 percentage point increase in the likelihood of receiving PANES for eligible households).

The coefficient of 0.129 suggests households eligible for PANES had an average 12.9 percentage point higher support for the government in 2007, during the program.

The coefficient of 0.118 suggests that households eligible for PANES had 11.8 percentage points more support for the government in 2008, even after the program had ended.

\pagebreak


## **Question: Write down the specifications used in row 2 of columns 1,2 and 3 of table 1. **

**Answer:**

<center>
**Column 1**
</center>

$$Gov.Support_{07}=\beta_0+\beta_1Eligibility+\epsilon$$


<center>
**Column 2** 
</center>

$$Gov.Support_{07}=\beta_0+\beta_1Eligibility+\beta_2Score+\beta_3(Eligibility:Score)+\epsilon$$

<center>
**Column 3**
</center>

$$Gov.Support_{07}=\beta_0+\beta_1Eligibility+\beta_2Score+\beta_3Score^2+\beta_4(Eligibility:Score)+\beta_5(Eligibility:Score)^2+\epsilon$$



\pagebreak


## **Replicate the results reported in row 2 of Table 1 columns 1, 2 and 3. Explain the difference between these specifications and interpret their coefficients.**
Hint: the variables listed in the table above after newtreat are the controls you will want to include.

**Code:**


```{r,results='asis'}
model1 <- lm(h_89_scaled ~ newtreat, data = df)

# Column 2: Linear score control (Eligibility + Score + Interaction)
model2 <- lm(h_89_scaled ~ newtreat + ind_reest + newtreat:ind_reest, data = df)

# Column 3: Quadratic score control (Eligibility + Score + Score^2 + Interaction)
model3 <- lm(h_89_scaled ~ newtreat + ind_reest + I(ind_reest^2) + newtreat:ind_reest
+ newtreat:I(ind_reest^2), data = df)

# Display the results using stargazer
stargazer(model1, model2, model3,type = "latex")
```


\pagebreak


## **What is the point of including all of these specifications?**

**Answer:**

No controls gives a simple estimate of how eligibility affects government support in 2007, without accounting for the potential impact of income or other covariates.

A linear control for ind_reest (predicted income), along with an interaction term between newtreat and ind_reest. This helps to assess whether the effect of eligibility on government support varies across different levels of predicted income (linearly).

A quadratic term for ind_reest (ind_reest^2) and its interaction with newtreat. This allows the model to capture any non-linear effects of income on government support and investigate whether the relationship between eligibility and government support changes at different levels of income in a non-linear way.

\pagebreak


## **Using the coefficients estimated above, write out the function you would use to predict the probability a household supports the current government based on their predicted income score: **

**a) If they are eligible for the transfer using the results from column 1.**

$$Gov.Support_{07}=0.772+0.129(1)=0.901$$

**b) If they are not eligible for the transfer using the results from column 1.**

$$Gov.Support_{07}=0.772+0.129(0)=0.772$$

**c) If they are eligible for the transfer using the results from column 2.**

$$Gov.Support_{07}=0.772+0.129(1)-0.011(Score)-1.916(Score)$$

$$Gov.Support_{07}=0.901-1.927(Score)$$

**d) If they are not eligible for the transfer using the results from column 2.**

$$Gov.Support_{07}=0.772+0.129(0)-0.011(Score)-1.916(Score)$$
$$Gov.Support_{07}=0.772-0.011(Score)$$


**e) If they are eligible for the transfer using the results from column 3.**



$$Gov.Support_{07}=0.901-0.011(Score)-40.457(Score^2)$$

**f) If they are not eligible for the transfer using the results from column 3.**

 
$$Gov.Support_{07}=0.901-0.011(Score)-40.457(Score^2)$$




\pagebreak


## **How narrow is the "bandwidth" used by the authors. Why does this matter? Check that the results are robust to a narrower bandwidth. **

**Code:**

```{r}
cutoff <- 0

# Example of different bandwidths (in terms of score)
bandwidths <- c(0.01, 0.05, 0.15)  # 1%, 5%, 10% bandwidths around the cutoff

# Subset data for each bandwidth
narrow_bandwidth_data <- lapply(bandwidths, function(bw) {
  subset(df, abs(df$ind_reest - cutoff) <= bw)
})


```

```{r}
models <- lapply(narrow_bandwidth_data, function(sub_data) {
  lm(h_89_scaled ~ newtreat * ind_reest + I(ind_reest^2), data = sub_data)
})
```

```{r,results='asis'}
stargazer(models[[1]], models[[2]], models[[3]],
          type = "latex",
          title = "Regression Results for Different Bandwidths",
          dep.var.labels = "Government Support 2007",
          digits = 3,
          model.names = FALSE,
          omit.stat = c("f", "ser")) 
```

**Answer:**

The author used 2% bandwidth.  All of the coefficients are very similar indicating that the results are robust, because the smaller bandwidth estimates are very similar to the larger bandwidth results.

\pagebreak


## **The authors attribute these effects to the causal effect of receiving the government transfers. What is the implied assumption behind this interpretation?**

**Answer:**

In a regression discontinuity design is that the potential outcome for a unit (household, in this case) that receives the treatment (eligibility for PANES) is continuous around the threshold for eligibility. This means that, had the threshold been slightly different, the outcome for eligible individuals would have been similar to the outcome for ineligible individuals, except for the effect of the treatment.

The eligibility rule creates a "cutoff" where households just above or below the threshold are assumed to be similar in all respects except for their eligibility status. This allows the comparison between households just below and just above the threshold to be interpreted as a randomized experiment near the cutoff. 

\pagebreak


## **What evidence do they provide to support this assumption?**

**Answer:**

"Because the targeting rule was thus insulated from political 
considerations, and its implementation was remarkably strict, assignment to the pro
gram near the threshold is “as good as random.”
 Around 18 months following the start of the program, households with income 
scores in the neighborhood of the threshold were surveyed and asked a series of 
questions, including their support for the current government. A second similar fol
low-up survey took place in 2008 after the program had already ended. The quasi
random assignment of applicants in the neighborhood of the threshold allows us to 
circumvent the problems of reverse causality, endogenous political targeting, and 
omitted variables highlighted above, and thus credibly estimate the impact of trans
fers on support for the incumbent. "



## **Was this threshold eligibility score specifically designed for this particular program? Why does this matter?**

**Answer:**

"The eligibility thresholds varied across the 
country’s five main administrative regions in order to entitle the same proportion of poor households in each area 
to the program. The regions are: Montevideo, North (Artigas, Salto, Rivera); Center-North (Paysandú, Río Negro, 
Tacuarembó, Durazno, Treinta y Tres, Cerro Largo); Center-South (Soriano, Florida, Flores, Lavalleja, Rocha); 
and South (Colonia, San José, Canelones, Maldonado). Only households with predicted income scores below a 
predetermined threshold were assigned to program treatment."



\pagebreak







